An Improved Approximate-Bayesian Model-choice Method for Estimating Shared Evolutionary History

### 2014-02-19 manuscripts/dpp-msbayes/dpp-msbayes.tex
Changing first person to singular in dpp ms.

\begin{abstract}
    To understand the processes that generate biodiversity, it is
    important to account for large-scale processes that affect the evolutionary
    history of groups of co-distributed populations of organisms.
    Such events predict temporally clustered divergences times, a pattern
    that can be estimated using genetic data from co-distributed species.
    We introduce a new approximate-Bayesian model for comparative
    phylogeographical model-choice that estimates the temporal distribution of
    divergences across taxa from multi-locus DNA sequence data.
    Our model is an extension of that implemented in \msb.
    By reparameterizing the model, introducing more flexible priors on
    demographic and divergence-time parameters, and implementing a
    non-parametric Dirichlet-process prior over divergence models, we improve
    the robustness, accuracy, and power of the method for estimating shared
    evolutionary history across taxa.
    Our results demonstrate the improved performance of the new method is due
    to (1) more appropriate priors on divergence-time and demographic
    parameters that avoid prohibitively small marginal likelihoods for models
    with more divergence events,
    and (2) the Dirichlet-process providing a flexible prior on divergence
    histories that does not strongly disfavor models with intermediate numbers
    of divergence events.
    % Our results demonstrate that the tendency of the \msb model to spuriously
    % support models of clustered divergences is caused by a combination of (1)
    % uniform priors on nuisance parameters hindering the marginal likelihoods of
    % models with more divergence-time parameters, and (2) a prior on divergence
    % models that disfavors models with intermediate numbers of divergence-time
    % parameters.
    The new method yields more robust estimates of posterior uncertainty, and
    thus greatly reduces the tendency of the model to incorrectly estimate
    biogeographically interesting models with strong support.

>>>>

\begin{abstract}
    To understand the processes that generate biodiversity, it is
    important to account for large-scale processes that affect the evolutionary
    history of groups of co-distributed populations of organisms.
    Such events predict temporally clustered divergences times, a pattern
    that can be estimated using genetic data from co-distributed species.
    I introduce a new approximate-Bayesian model for comparative
    phylogeographical model-choice that estimates the temporal distribution of
    divergences across taxa from multi-locus DNA sequence data.
    The model is an extension of that implemented in \msb.
    By reparameterizing the model, introducing more flexible priors on
    demographic and divergence-time parameters, and implementing a
    non-parametric Dirichlet-process prior over divergence models, I improve
    the robustness, accuracy, and power of the method for estimating shared
    evolutionary history across taxa.
    My results demonstrate the improved performance of the new method is due
    to (1) more appropriate priors on divergence-time and demographic
    parameters that avoid prohibitively small marginal likelihoods for models
    with more divergence events,
    and (2) the Dirichlet-process providing a flexible prior on divergence
    histories that does not strongly disfavor models with intermediate numbers
    of divergence events.
    % The results demonstrate that the tendency of the \msb model to spuriously
    % support models of clustered divergences is caused by a combination of (1)
    % uniform priors on nuisance parameters hindering the marginal likelihoods of
    % models with more divergence-time parameters, and (2) a prior on divergence
    % models that disfavors models with intermediate numbers of divergence-time
    % parameters.
    The new method yields more robust estimates of posterior uncertainty, and
    thus greatly reduces the tendency of the model to incorrectly estimate
    biogeographically interesting models with strong support.
